The Hosting Environment of the Advanced Resource Connector middleware

### 2008-09-08 doc/hed/ARCHED_article.tex
More text into HED document

@@

Each MCC has an entry method \texttt{process} which is called with Message being processed. It then processes Message by modifying it or creating new Message. Then MCC calls entry method of next MCC in the chain. For information how messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.

@@

The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs. Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component.

@@

@@

@@

\texttt{Relevant classes: Arc::SecAttr}

@@

\texttt{Relevant classes: ArcSec::SecHander}

@@

The services are dynamically loaded on start up just like the MCCs. They almost identic to MCCs with an exception that they constitute the last link in the Message Chain. They are attached to the Chain in the same manner as other MCCs. But differently from MCC the \texttt{process} method of the Service does not pass Message to other components of the Chain. Instead they have to process incoming Messages and produce outgoing ones.

@@

For reporting errors each \texttt{process} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.

@@

\texttt{Relevant classes: Arc::Loader}

@@

In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done in Python language.

% TODO it

@@

The server TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs. As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain. One could envision other MCCs having these properties but producing messages from other sources such as e.g. unix sockets.

@@

\ifthenelse{\boolean{verbtechdesc}}{
This MCC can be configuered with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}. The \texttt{<tcp:Port>} element is mandatory and should contain an integer corresponding to the TCP port to listen to. The \texttt{<tcp:Interface>} element is optional and is meant to identify the network interface to bind to. It is currently not used. The \texttt{<tcp:Version>} element is used to specify IP version. It is optional and should if present contain the single digit 4 or 6.
}{}

The server and client side MCCs are called \texttt{tcp.service} and \texttt{tcp.client} correspondingly. They are located in the shared library called \texttt{mcctcp}.

\subsubsection{TLS MCC}
The server TLS MCC enables ...

\ifthenelse{\boolean{brieftechdesc}}{
This MCC needs to be configured with a host certificate as well as trusted CA certificates.
}{}

\ifthenelse{\boolean{verbtechdesc}}{
The
}{}

@@

@@

@@

>>>>

\texttt{Relevant classes: Arc::MCC, Arc::Loader}

@@

Each MCC has an entry method \texttt{process()} which is called with Message being processed. It then processes Message by modifying it or creating new Message. Then MCC calls entry method of next MCC in the chain. For information how messages are handled and about memory management policies please see API description of MCC class in \cite{hed-api}.

@@

The MCC may implement some routing algorithm which means one MCC may have connections to multiple other MCCs. Typical scenario is that the HTTP MCC at the server side routes the HTTP messages with POST HTTP operation to a SOAP MCC but the messages with GET operation to for example a simple HTTPD service component. For that purpose <next> elements in MCC configuraton may have optional \texttt{id} attribute which allows to assign labels to all chain links to next MCCs in the chain. Supported labels are MCC dependant. By default simple MCCs support only one unnamed link. The MCCs with routing capabilities must have all supported labels documented.

@@

\texttt{Relevant classes: Arc::Payload, Arc::PayloadRawInterface, Arc::PayloadRaw, Arc::PayloadStreamInterface, Arc::PayloadStream, Arc::PayloadSOAP}

@@

\texttt{Relevant classes: Arc::MessageAttributes, Arc::Message}

@@

\texttt{Relevant classes: Arc::SecAttr, Arc::MessageAuth, Arc::MessageAuthContext, Arc::Message}

@@

\texttt{Relevant classes: ArcSec::SecHander, Arc::MCC}

@@

The services are dynamically loaded on start up just like the MCCs. They almost identic to MCCs with an exception that they constitute the last link in the Message Chain. They are attached to the Chain in the same manner as other MCCs. But differently from MCC the \texttt{process()} method of the Service does not pass Message to other components of the Chain. Instead they have to process incoming Messages and produce outgoing ones.

@@

For reporting errors each \texttt{process()} method returns instance of MCC\_Status class. That object carries predefined set of common error codes.

@@

\texttt{Relevant classes: Arc::Loader, Arc::Config, Arc::LoaderFactory, Arc::MCCFactory, Arc::ServiceFactory, Arc::DMCFactory, Arc::SecHandlerFactory, Arc::PDPFactory, Arc::ACCFactory}

@@

In order to facilitate the developement of services, API bindings for languages other than C++ are provided and some service developement has already been done in Python language. Currently only available language bindings are Python and Java. And currently it is possible to write only SOAP Service modules in Python and Java due to multiple inheritance limitation.

% TODO more of it

@@

\texttt{Plugin names: tcp.service, tcp.client}

\texttt{Library name: \{lib\}mcctcp}

\texttt{SEcuruty handler queues: incoming, outgoing}

% TODO: attributes and security attributes

The server side TCP MCC in the HED is special in that it produces messages by listening on a network socket rather than passing on messages from other MCCs. As such it spawns new thread for every new connection to handle Messages and their responses throughout the Message Chain. One could envision other MCCs having these properties but producing messages from other sources such as e.g. unix sockets.

@@

\ifthenelse{\boolean{verbtechdesc}}{
This MCC can be configuered with one or more \texttt{<tcp:Listen>} elements which in turn contain the elements \texttt{<tcp:Port>}, \texttt{<tcp:Interface>} and \texttt{<tcp:Version>}. The \texttt{<tcp:Port>} element is mandatory and should contain an integer corresponding to the TCP port to listen to. The \texttt{<tcp:Interface>} element is optional and is meant to identify the network interface to bind to. It is currently not used. The \texttt{<tcp:Version>} element is used to specify IP version. It is optional and should if present contain the single digit 4 or 6.

The server side TCP MCC generates \texttt{PayloadStreamInterface} payload in the Message passed to next MCC whihc can be used for communicationg through open TCP channel. Currently it ignores any payload attached to the returned Message.
}{}

The client side TCP MCC performs TCP connection to host and port specified in <Host> and <Port> elements inside <Connect> element of the MCC configuration. Then all incoming Messages of \texttt{process()} method are transfered ober TCP connection. Accepted Payload type of incoming Message is \texttt{PayloadRawInterface}. Returned Payload is of \texttt{PayloadStreamInterface} type. It represents established TCP conenction and \texttt{may} be used by previous MCCs in chain for direct communication. It is still prefered to call \texttt{process()} method instead.

\paragraph{Configuration schema}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xsd:schema
  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
  xmlns="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
  elementFormDefault="qualified">

    <xsd:simpleType name="Version_Type">
        <!-- This element defines TCP/IP protocol version.  -->
        <xsd:restriction base="xsd:string">
            <xsd:enumeration value="4"/>
            <xsd:enumeration value="6"/>
        </xsd:restriction>
    </xsd:simpleType>
    <xsd:element name="Version" type="Version_Type"/>

    <xsd:complexType name="Listen_Type">
        <!--
            This element defines listening TCP socket. If interface is missing socket
            is bound to all local interfaces (not supported). There may be multiple Listen elements.
        -->
        <xsd:sequence>
            <xsd:element name="Interface" type="xsd:string" minOccurs="0" maxOccurs="1"/>
            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
            <xsd:element name="Version" type="Version_Type" minOccurs="0" maxOccurs="1"/>
        </xsd:sequence>
    </xsd:complexType>
    <xsd:element name="Listen" type="Listen_Type"/>
    <xsd:complexType name="Connect_Type">
        <!--
            This element defines TCP connection to be established to specified Host at specified Port.
            If LocalPort is defined TCP socket will be bound to this port number (not supported).
        -->
        <xsd:sequence>
            <xsd:element name="Host" type="xsd:string" minOccurs="1" maxOccurs="1"/>
            <xsd:element name="Port" type="xsd:int" minOccurs="1" maxOccurs="1"/>
            <xsd:element name="LocalPort" type="xsd:int" minOccurs="0" maxOccurs="1"/>
        </xsd:sequence>
    </xsd:complexType>
    <xsd:element name="Connect" type="Connect_Type"/>
</xsd:schema>
\end{verbatim}

\subsection{TLS MCC}

\texttt{Plugin names: tls.service, tls.client}

\texttt{Library name: \{lib\}mcctls}

\texttt{Securuty handler queues: incoming, outgoing}

% TODO: attributes and security attributes

The server and client TLS MCCs provide transport level security (TLS) over any stream channel. Curently they interoperate well with TCP MCCs.

The server side MCC accepts payload of type PayloadStreamInterface. It then creates own instance of object inherited from PayloadStreamInterface bound to initial payload and passes it to next MCC. This object is maintained inside Message Context under name \texttt{tls.service} and is destroyed when Context becomes inactive. Currenlty this MCC does not expect any payload to be returned from rest of the chain and passes no payload to previous MCC.

The client side MCC behaves in similar way. It also establishes PayloadStreamInterface type object linked to same type of payload of next MCC. To obtain that last payload it makes a first call to next MCC with payload of type PayloadRawInterface and then uses returned payload - which is expected to be of PayloadStreamInterface type - to create own payload object with streaming capabilities and returns it back to previous MCC for further usage.

Both client ans server side MCCs are implemented using OpenSSL toolkit \cite{openssl}, use X.509 infrastructure \cite{x509} for establishing secure connection and may be configured to get private key, certfificate or proxy credentials from files esiding at local file system. It also possible to specify location of Certification Authority cerfificate or to use all certificates located in specified directory. for more information see configuration schema with comments below.

\paragraph{Configuration schema}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xsd:schema
  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
  xmlns="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
  xmlns:arc="http://www.nordugrid.org/schemas/ArcConfig/2007"
  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCTLS/2007"
  elementFormDefault="qualified">
    <xsd:complexType name="CACertificatesDir_Type">
      <xsd:simpleContent>
        <xsd:extension base="xsd:string">
          <xsd:attribute name="PolicyGlobus" type="xsd:boolean" use="optional" default="false"/>
        </xsd:extension>
      </xsd:simpleContent>
    </xsd:complexType>
    <!-- Location of private key.
      Default is /etc/grid-security/hostkey.pem for service
      and none for client. -->
    <xsd:element name="KeyPath" type="xsd:string"/>
    <!-- Content of private key - not supported -->
    <xsd:element name="Key" type="xsd:string"/>
    <!-- Location of public certificate.
      Default is /etc/grid-security/hostcert.pem for service
      and none for client. -->
    <xsd:element name="CertificatePath" type="xsd:string"/>
    <!-- Content of public certificate - not supported -->
    <xsd:element name="Certificate" type="xsd:string"/>
    <!-- Location of proxy credentials - includes certificates, key and
      chain of involved certificates. Overwrites elements Key, KeyPath,
      Certificate and CertificatePath.
      Default is none for client and none for service. -->
    <xsd:element name="ProxyPath" type="xsd:string"/>
    <!-- Content of proxy credentials - not supported -->
    <xsd:element name="Proxy" type="xsd:string"/>
    <!-- Location of certificate of CA. Default is none. -->
    <xsd:element name="CACertificatePath" type="xsd:string"/>
    <!-- Content of certificate of CA - not supported -->
    <xsd:element name="CACertificate" type="xsd:string"/>
    <!-- Directory containing certificates of accepted CAs.
      Default is /etc/grid-security/ . -->
    <xsd:element name="CACertificatesDir" type="xsd:string"/>
</xsd:schema>
\end{verbatim}

@@

\texttt{Plugin names: http.service, http.client}

\texttt{Library name: \{lib\}mcchttp}

\texttt{Securuty handler queues: incoming, outgoing}

% TODO: attributes and security attributes

The server side HTTP MCC accepts messages with \texttt{PayloadStreamInterface} payload and parses HTTP related information from it. Information from the HTTP header is added to the Message Attributes. The body of HTTP message is passed to next MCC as \texttt{PayloadRawInterface} payload. In response this MCC expects also the Message with \texttt{PayloadRawInterface}. It is then prepended with HTTP reponse header and pushes it into initially providded stream channel. for output it returns empty \texttt{PayloadRawInterface} payload.

This MCC routes results to multiple next MCCs in the chain. For that it accepts only labeled <next> elements in cinfoguration. Label names are those of HTTP methods (uppercase). HTTP messages will be routed to their destinations according HTTP method requested.

The client side HTTP MCC will accept \texttt{PayloadRawInterface} payload as HTTP body and after prepending it with HTTP information passes to next MCC also as \texttt{PayloadRawInterface}. It accepts \texttt{PayloadStreamInterface} in response and after processing passes \texttt{PayloadRawInterface} back trhough the chain.

\paragraph{Configuration schema}

\begin{verbatim}
<?xml version="1.0" encoding="UTF-8"?>
<xsd:schema
  xmlns:xsd="http://www.w3.org/2001/XMLSchema"
  xmlns="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
  xmlns:arc="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
  targetNamespace="http://www.nordugrid.org/schemas/ArcMCCHTTP/2007"
  elementFormDefault="qualified">

    <!--
        These elements define endpoint and HTTP method for client HTTP MCC.
    -->
    <xsd:element name="Endpoint" type="xsd:string"/>
    <xsd:element name="Method" type="xsd:string"/>
</xsd:schema>
\end{verbatim}

@@

\texttt{Plugin names: soap.service, soap.client}

\texttt{Library name: \{lib\}mccsoap}

\texttt{Securuty handler queues: incoming, outgoing}

% TODO: attributes and security attributes

These MCCs convert payloads between dat chunks presented by PayloadRawInterface type object into dedicated PayloadSOAP payloads and vice versa. Currently it has no specific configuration parameters.

@@

\subsection{Simple List Auth}

\texttt{Plugin name: arc.authz}

\texttt{Library name: \{lib\}arcpdc}

\subsection{Username Token}

\texttt{Plugin name: usernametoken.handler}

\texttt{Library name: \{lib\}arcpdc}

\subsection{X.509 Token}

\texttt{Plugin name: x509token.handler}

\texttt{Library name: \{lib\}arcpdc}
